---
title: "Task 3 - Santa Barbara Urchins"
author: "Meghna Rao"
date: "10/29/2021"
output: html_document
---
# Overview

This report is exploring data on red and purple urchins observed at 5 sites along the Santa Barbara coast.

**Data Citation:** [Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative.](https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b.)

# Setup

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# Attach Packages
library(tidyverse)
library(here)
library(janitor)
```

# Read in the Data

Convert the column names into lower snake case.

```{r}
# Read in Data
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

# Section 1 - Size Comparisons of Purple and Red Urchins

## Data Visualization

Create a faceted histogram to compare the size distribution between purple and red urchins.

```{r}
urchins %>% 
  ggplot(aes(x = size)) + 
  geom_histogram(bins = 10, color = "black") +
  facet_wrap(~ common_name) +
  labs(x = "Urchin Diameter (cm)",
       y = "Number of Sea Urchins")
```

